---
title: "01_Import_to_Phyloseq"
output: html_document
date: "2025-02-10"
---
---
title: "Rock3_lulu_2024-12-03"
output: html_document
date: "2024-12-03"
---
```{r}
library(lulu)
library(tidyverse)
library(ggplot2)
library(readxl)
library(metagMisc)
library(phyloseq)
set.seed(110011)
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
```


#load asv tables, reformat for phyloseq
```{r run1}
MiSeqReads_r1 <- read.csv("/Users/stephaniematthews/Documents/Rockfish/Processed_MiSeq/MDL_2024-04-13/outputs/2024-04-13_MDL_ASV_table.csv")
MiSeqReads_r1_wide <- MiSeqReads_r1 %>% pivot_wider(names_from = Sample_name, values_from = nReads, id_cols = Hash, values_fill = 0)

MiSeqReads_r1_wide <- as.data.frame(MiSeqReads_r1_wide)
rownames(MiSeqReads_r1_wide) <- MiSeqReads_r1_wide$Hash
MiSeqReads_r1_wide_nohash <- MiSeqReads_r1_wide[,-1]
MiSeqReads_r1_wide_nohash <- as.data.frame(lapply(MiSeqReads_r1_wide_nohash, as.numeric))
rownames(MiSeqReads_r1_wide_nohash) <- MiSeqReads_r1_wide$Hash
colnames(MiSeqReads_r1_wide_nohash) <- str_split_i(colnames(MiSeqReads_r1_wide_nohash), pattern = "_", 1)
```

```{r run2}
MiSeqReads_r2 <- read.csv("/Users/stephaniematthews/Documents/Rockfish/Processed_MiSeq/MDL_2024-07-19_Rockfish2/outputs/2024-07-19_Rockfish2_MDL_ASV_table.csv")
MiSeqReads_r2_wide <- MiSeqReads_r2 %>% pivot_wider(names_from = Sample_name, values_from = nReads, id_cols = Hash, values_fill = 0)

MiSeqReads_r2_wide <- as.data.frame(MiSeqReads_r2_wide)
rownames(MiSeqReads_r2_wide) <- MiSeqReads_r2_wide$Hash
MiSeqReads_r2_wide_nohash <- MiSeqReads_r2_wide[,-1]
MiSeqReads_r2_wide_nohash <- as.data.frame(lapply(MiSeqReads_r2_wide_nohash, as.numeric))
rownames(MiSeqReads_r2_wide_nohash) <- MiSeqReads_r2_wide$Hash
```

```{r run misfits}
MiSeqReads_misfits <- read.csv("/Users/stephaniematthews/Documents/Rockfish/Processed_MiSeq/MDL_2024-10-11_Misfits/outputs/2024-10-11_Misfits_MDL_ASV_table.csv")
MiSeqReads_misfits_wide <- MiSeqReads_misfits %>% pivot_wider(names_from = Sample_name, values_from = nReads, id_cols = Hash, values_fill = 0)

MiSeqReads_misfits_wide <- as.data.frame(MiSeqReads_misfits_wide)
rownames(MiSeqReads_misfits_wide) <- MiSeqReads_misfits_wide$Hash
MiSeqReads_misfits_wide_nohash <- MiSeqReads_misfits_wide[,-1]
MiSeqReads_misfits_wide_nohash <- as.data.frame(lapply(MiSeqReads_misfits_wide_nohash, as.numeric))
rownames(MiSeqReads_misfits_wide_nohash) <- MiSeqReads_misfits_wide$Hash
colnames(MiSeqReads_misfits_wide_nohash) <- str_split_i(colnames(MiSeqReads_misfits_wide_nohash), pattern = "_", 1)

```

```{r run3}
MiSeqReads_r3 <- read.csv("/Users/stephaniematthews/Documents/Rockfish/Processed_MiSeq/MDL_2024-12-02_Rockfish3/outputs/2024-12-02_Rockfish3_MDL_ASV_table.csv")
MiSeqReads_r3_wide <- MiSeqReads_r3 %>% pivot_wider(names_from = Sample_name, values_from = nReads, id_cols = Hash, values_fill = 0)

MiSeqReads_r3_wide <- as.data.frame(MiSeqReads_r3_wide)
rownames(MiSeqReads_r3_wide) <- MiSeqReads_r3_wide$Hash
MiSeqReads_r3_wide_nohash <- MiSeqReads_r3_wide[,-1]
MiSeqReads_r3_wide_nohash <- as.data.frame(lapply(MiSeqReads_r3_wide_nohash, as.numeric))
rownames(MiSeqReads_r3_wide_nohash) <- MiSeqReads_r3_wide$Hash
colnames(MiSeqReads_r3_wide_nohash) <- str_split_i(colnames(MiSeqReads_r3_wide_nohash), pattern = "_", 1)

```


#load metadata, reformat for phyloseq
```{r run1}
samples1 <- read.delim("/Users/stephaniematthews/Documents/Rockfish/MiSeq_analysis/metadata_miseqs/MDL_2024-04-20_MegsRun1_metadata.csv", sep = ",")
rownames(samples1) <- paste0("MDL.", samples1$X)

```

```{r run2}
samples2 <- read.delim("/Users/stephaniematthews/Documents/Rockfish/MiSeq_analysis/metadata_miseqs/MDL_2024-07-19_Rockfish2_metadata.csv", sep = ",")
rownames(samples2) <- samples2$Sample_name
rownames(samples2) <- gsub("-", ".", rownames(samples2))
samples2$Type[samples2$Type == "helvomaculatus"] <- "helvomaculatus.tissue"
samples2$Type[samples2$Type == "ruberrimus"] <- "ruberrimus.tissue"

```

```{r misfits}
samplesmisfits <- read.delim("/Users/stephaniematthews/Documents/Rockfish/MiSeq_analysis/metadata_miseqs/MDL_2024-09-10_Misfits_metadata.csv", sep = ",")
rownames(samplesmisfits) <- samplesmisfits$Sample_ID
rownames(samplesmisfits) <- gsub("-", ".", rownames(samplesmisfits))
```

```{r run3}
samples3 <- read.delim("/Users/stephaniematthews/Documents/Rockfish/MiSeq_analysis/metadata_miseqs/MDL_2024-10-07_Rockfish3_metadata.csv", sep = ",")
rownames(samples3) <- samples3$Sample_name
rownames(samples3) <- gsub("-", ".", rownames(samples3))

```


#load fastas, reformat for phyloseq
```{r run1}
bs1 <- Biostrings::readDNAStringSet("/Users/stephaniematthews/Documents/Rockfish/Processed_MiSeq/MDL_2024-04-13/outputs/2024-04-13_MDL_hash_key.fasta")
```

```{r run2}
bs2 <- Biostrings::readDNAStringSet("/Users/stephaniematthews/Documents/Rockfish/Processed_MiSeq/MDL_2024-07-19_Rockfish2/outputs/2024-07-19_Rockfish2_MDL_hash_key.fasta")
```

```{r misfits}
bsmisfits <- Biostrings::readDNAStringSet("/Users/stephaniematthews/Documents/Rockfish/Processed_MiSeq/MDL_2024-10-11_Misfits/outputs/2024-10-11_Misfits_MDL_hash_key.fasta")
```

```{r run3}
bs3 <- Biostrings::readDNAStringSet("/Users/stephaniematthews/Documents/Rockfish/Processed_MiSeq/MDL_2024-12-02_Rockfish3/outputs/2024-12-02_Rockfish3_MDL_hash_key.fasta")
```


#convert for phyloseq, merge phyloseq pieces
```{r}
#asv table 1
MiSeqReads_r1_wide_nohash_ps <- otu_table(MiSeqReads_r1_wide_nohash, taxa_are_rows = T)
taxa_names(MiSeqReads_r1_wide_nohash_ps) <- rownames(MiSeqReads_r1_wide_nohash)

#asv table 2
MiSeqReads_r2_wide_nohash_ps <- otu_table(MiSeqReads_r2_wide_nohash, taxa_are_rows = T)
taxa_names(MiSeqReads_r2_wide_nohash_ps) <- rownames(MiSeqReads_r2_wide_nohash)

#asv table misfits
MiSeqReads_misfits_wide_nohash_ps <- otu_table(MiSeqReads_misfits_wide_nohash, taxa_are_rows = T)
taxa_names(MiSeqReads_misfits_wide_nohash_ps) <- rownames(MiSeqReads_misfits_wide_nohash)

#asv table 3
MiSeqReads_r3_wide_nohash_ps <- otu_table(MiSeqReads_r3_wide_nohash, taxa_are_rows = T)
taxa_names(MiSeqReads_r3_wide_nohash_ps) <- rownames(MiSeqReads_r3_wide_nohash)

#metadata 1
MiSeq_metadata1_ps <- sample_data(samples1)

#metadata 2
MiSeq_metadata2_ps <- sample_data(samples2)

#metadata misfits
MiSeq_metadatamisfits_ps <- sample_data(samplesmisfits)

#metadata 3
MiSeq_metadata3_ps <- sample_data(samples3)


#merge
ps_r1 <- merge_phyloseq(MiSeqReads_r1_wide_nohash_ps, bs1, MiSeq_metadata1_ps)
ps_r2 <- merge_phyloseq(MiSeqReads_r2_wide_nohash_ps, bs2, MiSeq_metadata2_ps)
ps_misfits <- merge_phyloseq(MiSeqReads_misfits_wide_nohash_ps, bsmisfits, MiSeq_metadatamisfits_ps)
ps_r3 <- merge_phyloseq(MiSeqReads_r3_wide_nohash_ps, bs3, MiSeq_metadata3_ps)

ps_MDL <- merge_phyloseq(ps_r1, ps_r2, ps_misfits, ps_r3)

ps_MDL %>%
      refseq() %>%
      Biostrings::writeXStringSet("~/Documents/Rockfish/lulu/fasta_by_run/merged_four.fasta", append=FALSE,
                                  compress=FALSE, compression_level=NA, format="fasta")


#ps_MDL <- subset_samples(ps_MDL, Site %in% c("Possession", "Mulkiteo", "HC-8", "HC-9", "HC-1", "AI-8", "AI-2", "HC-0", "HC-4", "AI-1", "HC-6", "HC-5"))

```

##On linux, blast all ASV sequences against each other
makeblastdb -in merged_four.fasta -parse_seqids -dbtype nucl


blastn -db merged_four.fasta -outfmt '6 qseqid sseqid pident' -out match_list_four.txt -qcov_hsp_perc 80
-perc_identity 84 -query merged_four.fasta

#run lulu
```{r}


##Make the combined ASV table because Tobias suggested it works best if you do it with all samples at once.
  
ASV_wide <- data.frame(otu_table(ps_MDL, taxa_sums(ps_MDL)))

ASV_wide <- ASV_wide[rowSums(ASV_wide) > 0,]
  
##input blast matchlist to R
matchlist <- read.table("/Users/stephaniematthews/Documents/Rockfish/lulu/fasta_by_run/match_list_four.txt", header=FALSE,as.is=TRUE, stringsAsFactors=FALSE)

## run lulu
curated_result <- lulu(ASV_wide, matchlist, minimum_ratio_type = "avg", minimum_ratio = 1, minimum_match = 84, minimum_relative_cooccurence = 0.95)
curated_result$discarded_count
curated_result$curated_count
lulu_filtered_ASVs <- curated_result$curated_table
good_taxa_fourdatasets <- rownames(lulu_filtered_ASVs)

```

## insect taxonomy 
```{r}
trained_classifier_20241011_tuna_outgroups_min70 <- readRDS("~/Documents/Rockfish/AnnotationDatabases_google_22April/steph_taxonomy_approaches/Dloop-Insect/trained_classifier_20241011_tuna_outgroups_min70.rds")

stephsclassifier <- trained_classifier_20241011_tuna_outgroups_min70

extendmax <- function(n){
  a <- attributes(n)
  attr(n, 'maxlength') <- a$maxlength+20
  n
}
extendmin <- function(n){
  a <- attributes(n)
  attr(n, 'minlength') <- a$minlength-20
  n
}

stephsclassifiermod <- dendrapply(stephsclassifier, extendmax)
stephsclassifiermod <- dendrapply(stephsclassifiermod, extendmin)

all_rockfish_4runs <- ape::read.dna("~/Documents/Rockfish/lulu/fasta_by_run/merged_four.fasta", format = "fasta")
good_rockfish_4runs <- all_rockfish_4runs[names(all_rockfish_4runs) %in% good_taxa_fourdatasets]

classified_steph2 <- insect::classify(good_rockfish_4runs, stephsclassifiermod, threshold = 0.7, metadata = TRUE, ping = 1, mincount = 1, decay = FALSE, species = 'all', offset = -40)
write.csv(classified_steph2, file = "/Users/stephaniematthews/Documents/Rockfish/Conservation_Letters/Data/classified_4runs_mersseqs_insect_70threshold.csv")

```

##reformat insect taxonomy
```{r}
MiSeqTaxonomy <- read.csv("/Users/stephaniematthews/Documents/Rockfish/Conservation_Letters/Data/classified_4runs_mersseqs_insect_70threshold.csv")
MiSeqTaxonomy_phyloseq <- MiSeqTaxonomy[c("kingdom", "phylum", "class", "order", "family", "genus", "species", "score", "NNtaxon", "representative")]

colnames(MiSeqTaxonomy_phyloseq) <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species", "score", "nntaxon", "representative")
MiSeqTaxonomy_phyloseq <- as.data.frame(MiSeqTaxonomy_phyloseq)
rownames(MiSeqTaxonomy_phyloseq) <- MiSeqTaxonomy$representative

MiSeqTaxonomy_phyloseq$Phylum[(MiSeqTaxonomy_phyloseq$Phylum == "")] <- MiSeqTaxonomy_phyloseq$Kingdom[(MiSeqTaxonomy_phyloseq$Phylum == "")]
MiSeqTaxonomy_phyloseq$Class[(MiSeqTaxonomy_phyloseq$Class == "")] <- MiSeqTaxonomy_phyloseq$Phylum[(MiSeqTaxonomy_phyloseq$Class == "")]
MiSeqTaxonomy_phyloseq$Order[(MiSeqTaxonomy_phyloseq$Order == "")] <- MiSeqTaxonomy_phyloseq$Class[(MiSeqTaxonomy_phyloseq$Order == "")]
MiSeqTaxonomy_phyloseq$Family[(MiSeqTaxonomy_phyloseq$Family == "")] <- MiSeqTaxonomy_phyloseq$Order[(MiSeqTaxonomy_phyloseq$Family == "")]
MiSeqTaxonomy_phyloseq$Genus[(MiSeqTaxonomy_phyloseq$Genus == "")] <- MiSeqTaxonomy_phyloseq$Family[(MiSeqTaxonomy_phyloseq$Genus == "")]
MiSeqTaxonomy_phyloseq$Species[(MiSeqTaxonomy_phyloseq$Species == "")] <- MiSeqTaxonomy_phyloseq$Genus[(MiSeqTaxonomy_phyloseq$Species == "")]
MiSeqTaxonomy_phyloseq$Species[(MiSeqTaxonomy_phyloseq$Species == "Sebastes sinensis")] <- "Sebastes diploproa"

commonnames <- read_excel("/Users/stephaniematthews/Documents/Rockfish/Sebastes_flock/20241030_names.xlsx", )
MiSeqTaxonomy_phyloseq_insect <- dplyr::left_join(MiSeqTaxonomy_phyloseq, commonnames, by = join_by(Species == GenusSpecies)) %>% select(c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species", "G.speciesC", "score", "nntaxon", "representative"))
MiSeqTaxonomy_phyloseq_insect$G.speciesC[is.na(MiSeqTaxonomy_phyloseq_insect$G.speciesC)] <- MiSeqTaxonomy_phyloseq_insect$Species[is.na(MiSeqTaxonomy_phyloseq_insect$G.speciesC)]
colnames(MiSeqTaxonomy_phyloseq_insect) <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Binomial", "Species", "score", "nntaxon", "representative")

rownames(MiSeqTaxonomy_phyloseq_insect) <- MiSeqTaxonomy_phyloseq_insect$representative
```


#merge into final cleaned phyloseq
```{r}
#asv table
curated_table <- otu_table(curated_result$curated_table, taxa_are_rows = T)
taxa_names(curated_table) <- rownames(curated_result$curated_table)
sample_names(curated_table) <- str_split_i(sample_names(curated_table), pattern = "_", 1)

#tax table
curated_taxonomy <- tax_table(as.matrix(MiSeqTaxonomy_phyloseq_insect))
taxa_names(curated_taxonomy) <- rownames(curated_taxonomy)

#fasta seqs
curated_seqs <- Biostrings::readDNAStringSet("~/Documents/Rockfish/lulu/fasta_by_run/merged_four.fasta")

#sample metadata
curated_metadata <- sample_data(ps_MDL)
sample_names(curated_metadata) <- str_split_i(sample_names(curated_metadata), pattern = "_", 1)

#merge
curated_MDL <- merge_phyloseq(curated_table, curated_taxonomy, curated_metadata, curated_seqs)

sample_names(curated_MDL) <- gsub("MeredithDloop.", "", sample_names(curated_MDL))
sample_names(curated_MDL) <- gsub("MDL", "", sample_names(curated_MDL))

rownames(samples2) <- gsub("-", ".", rownames(samples2))

curated_MDL_sebastes <- subset_taxa(curated_MDL, Genus == "Sebastes")
sample_data(curated_MDL_sebastes)$Sebastesreads <- sample_sums(curated_MDL_sebastes)
sample_data(curated_MDL_sebastes)$prop_sebastes <- sample_data(curated_MDL_sebastes)$Sebastesreads/sample_data(curated_MDL_sebastes)$readswithprimers



sample_data(curated_MDL_sebastes)$Sample_name <- gsub("MeredithDloop.", "", sample_data(curated_MDL_sebastes)$Sample_name)
sample_data(curated_MDL_sebastes)$Sample_name <- gsub("MDL", "", sample_data(curated_MDL_sebastes)$Sample_name)
sample_data(curated_MDL_sebastes)$Sample_name <- str_split_i(sample_data(curated_MDL_sebastes)$Sample_name, pattern = "_", 1)
curated_MDL_sebastes <- subset_samples(curated_MDL_sebastes, Sample_name != "24A")
sample_data(curated_MDL_sebastes)$Observed <- estimate_richness(curated_MDL_sebastes, measures = "Observed")$Observed

field_curated_MDL <- subset_samples(curated_MDL_sebastes, Basin %in% c("HC", "AI"))
field_curated_MDL <- subset_samples(field_curated_MDL, is.na(Location))
field_curated_MDL <- subset_samples(field_curated_MDL, Sample_name != "MeredithDloop-31A_S55")
field_curated_MDL <- subset_taxa(field_curated_MDL, taxa_sums(field_curated_MDL) > 0)

field_curated_MDL_abundant <- subset_samples(field_curated_MDL, sample_sums(field_curated_MDL) > 500)
field_curated_MDL_abundant <- subset_taxa(field_curated_MDL_abundant, taxa_sums(field_curated_MDL_abundant) > 0)


save(field_curated_MDL, file = "/Users/stephaniematthews/Documents/Rockfish/Conservation_Letters/Data/Rockfish_phyloseq.RData")

field_curated_MDL %>%
      refseq() %>%
      Biostrings::writeXStringSet("/Users/stephaniematthews/Documents/Rockfish/Conservation_Letters/Data/Rockfish_phyloseq.fna", append=FALSE,
                                  compress=FALSE, compression_level=NA, format="fasta")
#rm(list = ls())
load("/Users/stephaniematthews/Documents/Rockfish/Conservation_Letters/Data/Rockfish_phyloseq.RData")
```



